\section{Evaluation}\seclabel{Evaluation}
In this section, we describe and evaluate Lucy: a prototype implementation of
our decision procedures and system models. A more complete evaluation can be
found in \cite{whittaker2018interactive}.
%
Lucy includes a Python implementation of the interactive decision procedures
described in \algoref{InteractiveDecisionProcedure} and
\algoref{ArbitraryStartInteractiveDecisionProcedure}. Users
specify objects, transactions, invariants, and segmentations in Python.
%
Lucy also includes a C++ implementation of the \invariantconfluence{} and
segmented \invariantconfluence{} system models.

We now evaluate the practicality and efficiency of our decision procedure
prototypes. Specifically, we show that the user burden of specifying objects
and interactively checking them for invariant confluence is not too onerous.

\example[Foreign Keys]\examplelabel{ForeignKeysEval}
A 2P-Set $X = (A_X, R_X)$ is a set CRDT composed of a set of additions $A_X$
and a set of removals $R_X$~\cite{shapiro2011comprehensive}. We view the state
of the set $X$ as the difference $A_X - R_X$ of the addition and removal sets.
To add an element $x$ to the set, we add $x$ to $A_X$. Similarly, to remove $x$
from the set, we add it to $R_X$. The merge of two 2P-sets is a pairwise union
(i.e. $(A_X, R_X) \join (A_Y, R_Y) = (A_X \cup A_Y, R_X \cup R_Y)$).

We can use 2P-sets to model a simple relational database with foreign key
constraints. Let object $O = (X, Y) = ((A_X, R_X), (A_Y, R_Y))$ consist of a
pair of two 2P-Sets $X$ and $Y$, which we view as relations. Our invariant $X
\subseteq Y$ (i.e. $(A_X - R_X) \subseteq (A_Y - R_Y)$) models a foreign key
constraint from $X$ to $Y$. We ran our decision procedure on the object with
initial state $((\emptyset, \emptyset), (\emptyset, \emptyset))$ and with
transactions that allow arbitrary insertions and deletions into $X$ and $Y$.
After less than a tenth of a second, the decision procedure produced a
reachable counterexample witnessing the fact that the object is not
\invariantconfluent{}. A concurrent insertion into $X$ and deletion from $Y$
can lead to a state that violates the invariant. This object is not
\invariantconfluent{} and therefore not \invariantclosed{}.
Thus, existing systems that depend on \invariantclosure{} as a sufficient
condition are unable to conclude definitively that the object is \emph{not}
\invariantconfluent{}.

We also reran the decision procedure, but this time with insertions into $X$
and deletions from $Y$ disallowed. In less than a tenth of a second, the
decision procedure correctly deduced that the object is now
\invariantconfluent{}. These results were manually proven
in~\cite{bailis2014coordination}, but our tool was able to confirm them
automatically in a negligible amount of time.

\example[Escrow Transactions]\examplelabel{EscrowTransactionsEval}
Escrow transactions are a concurrency control technique that allows a database
to execute transactions that increment and decrement numeric values with more
concurrency than is otherwise possible with general-purpose techniques like
two-phase locking~\cite{o1986escrow}. The main idea is that a portion of the
numeric value is put in escrow, after which a transaction can freely decrement
the value so long as it is not decremented by more than the amount that has
been escrowed. We show how segmented \invariantconfluence{} can be used to
implement escrow transactions.

Consider again the PN-Counter $s = (p_1, p_2, p_3), (n_1, n_2, n_3)$ from
\exampleref{CounreachableExample} replicated on three servers with transactions
to increment and decrement the PN-Counter. In
\exampleref{CounreachableExample}, we found that concurrent decrements violate
\invariantconfluence{} which led us to a segmentation which prohibited
concurrent decrements. We now propose a new segmentation with escrow amount $k$
that will allow us to perform concurrent decrements that respect the escrowed
value. The first segment $(\setst{(p_1, p_2, p_3), (n_1, n_2, n_3)}{p_1, p_2,
p_3 \geq k \land n_1, n_2, n_3 \leq k}, T)$ allows for concurrent increments
and decrements so long as every $p_i \geq k$ and every $n_i \leq k$.
Intuitively, this segment represents the situation in which every server has
escrowed a value of $k$. Each server can decrement freely, so long as they don't
exceed their escrow budget of $k$. The second segment is the one presented in
\exampleref{CounreachableExample} which prohibits concurrent decrements. We ran
our decision procedure on this example and it concluded that it was segmented
\invariantconfluent{} in less than a tenth of a second and without any human
interaction.

\vspace{4pt}
\textbf{Further Evaluation.}
In~\cite{whittaker2018interactive}, we also specify workloads involving
\exampleref{Ints}, an auction application, and TPC-C. Lucy processes all of
these workloads, as well as the workloads described above, in less than half a
second, and no workload requires more than 66 lines of Python code to specify.
This shows that the user burden of specification is not too high and that our
decision procedures are efficient enough for interactive use.

In~\cite{whittaker2018interactive}, we also evaluate the performance of our
segmented \invariantconfluence{} system model.
%
Namely, we show that segmented \invariantconfluent{} replication can achieve an
order of magnitude higher throughput compared to linearizable replication, but
the throughput improvements decrease as we increase the fraction of
transactions that require coordination. For example, with 5\% coordinating
transactions, segmented \invariantconfluent{} replication performs over an
order of magnitude better than linearizable replication; with 50\%, it performs
as well; and with 100\%, it performs two times worse.
%
% We also measure the scalability of segmented \invariantconfluent{} replication.
% Expectedly, low coordination workloads scale well, but increasing coordination
% decreases scalability.
