\section{Sufficient Conditions}
In this section, we present sufficient (but not always necessary) conditions
for state-based and operation-based invariant-confluence. We also explain
circumstances under which the sufficient conditions are also necessary.

\begin{definition}
  We say a state-based object $O = (S, \join)$ is \defword{\Iclosed{}} if for
  all states $s_1$ and $s_2$, if $I(s_1)$ and $I(s_2)$, then $I(s_1 \join
  s_2)$.
  \[
    I(s_1) \land I(s_2) \implies I(s_1 \join s_2)
  \]
  Similarly, we say on operation-based object $O = S$ is \defword{\Iclosed{}}
  if for all states $s_1$ and $s_2$ and all transactions $t$, if $I(s_1)$,
  $I(s_2)$, and $I(t(s_1)(s_1))$, then $I(t(s_1)(s_2))$.
  \[
    I(s_1) \land I(s_2) \land I(t(s_1)(s_1)) \implies I(t(s_1)(s_2))
  \]
\end{definition}

\begin{claim}\clmlabel{StateBasedSufficient}
  Given a state-based object $O = (S, \join)$, a start state $s_0$, a set of
  transactions $T$, and an invariant $I$, if $O$ is \Iclosed{} and $I(s_0)$,
  then $O$ is \sTIconfluent.
\end{claim}
\begin{elidableproof}
  Let $e_1, e_2$ be $\sTIreachable{}$ expressions such that $\Irec{e_1}$ and
  $\Irec{e_2}$. Let $s_1 = eval(e_1)$ and $s_2 = eval(e_2)$. Then, $I(s_1)$ and
  $I(s_2)$. Because $O$ is \Iclosed{}, $I(s_1 \join s_2)$, so $I(e_1 \join
  e_2)$. Thus, by \clmref{StateBasedTwoIconfluenceDefs}, $O$ is
  \sTIconfluent{}.
\end{elidableproof}

\begin{claim}\clmlabel{OpBasedSufficient}
  Given a operation-based object $O = S$, a start state $s_0$, a set of
  transactions $T$, and an invariant $I$, if $O$ is \Iclosed{} and $I(s_0)$,
  then $O$ is \sTIconfluent.
\end{claim}
\begin{elidableproof}
  Let $t \in T$ and $e_1, e_2$ be $\sTIreachable{}$ expressions such that
  $\Irec{e_1}$, $\Irec{e_2}$, and $I(t(e_1)(e_1))$. Let $s_1 = eval(e_1)$ and
  $s_2 = eval(s_2)$. Then, $I(s_1)$, $I(s_2)$, $I(t(s_1)(s_1))$, so
  $I(t(s_1)(s_2))$ because $O$ is \Iclosed. Thus, $I(t(e_1)(e_2))$, so by
  \clmref{OpBasedTwoIconfluenceDefs}, $O$ is \sTIconfluent.
\end{elidableproof}

As a special case, \clmref{StateBasedSufficient} tells us that a state-based
object $O$ is \sTIconfluent{} if $I$ is monotonic and $O$ is a semilattice.

\begin{claim}\clmlabel{Monotonicity}
  Let $O = (S, \join)$ be a semilattice. Let $s \leq s' \defeq s \join s' =
  s'$. Let $s_0$ be an arbitrary start state. Let $T$ be an arbitrary set of
  transactions. Let invariant $I$ be \defword{monotonic}: for all $s, s'\in S$,
  if $s \leq s'$, then $I(s) \implies I(s')$. Also assume $I(s_0)$. Then, $T$
  is \sTIconfluent{}.
\end{claim}
\begin{elidableproof}
  Let $s_1$ and $s_2$ be arbitrary states such that $I(s_1)$ and $I(s_2)$. $s_1
  \leq s_1 \join s_2$, so by monotonicity, $I(s_1 \join s_2)$. Thus, by
  \clmref{StateBasedSufficient}, $O$ is \sTIconfluent{}.
\end{elidableproof}

Unfortunately, the converses of \clmref{StateBasedSufficient} and
\clmref{OpBasedSufficient} are not true. Being \sTIconfluent{} does not
necessarily imply being \Iclosed{}, as the next two examples show.

\begin{example}\examplelabel{StateBasedNotNecessary}
  Consider the state-based object $O = (S, \join) = (\ints, \max) \times
  (\ints, \max)$ and start state $s_0 = (0, 0)$. Let $T$ consist of two
  transactions $t_{x+1} \defeq \lambda (x, y).\ (x + 1, y)$ and $t_{y-1} =
  \lambda (x, y).\ (x, y - 1)$. Let $I = \setst{(x, y)}{xy \leq 0}$.

  $\setst{(x, y)}{\sTIreachable{(x, y)}} = \nats \times -\nats \subseteq I$, so
  $O$ is \sTIconfluent. However, let $s_1 = (-1, 1)$ and $s_2 = (1, -1)$.
  $I(s_1)$ and $I(s_2)$. However, $s_1 \join s_2 = (1, 1)$, and $\lnot I((1,
  1))$. Thus, $O$ is not \Iclosed{}.
\end{example}

\begin{example}
  Consider the operation-based object $O = S = \ints \times \ints$ and start
  state $s_0 = (0, 0)$. Let $T$ consist of two transactions $t_{x+1} \defeq
  \lambda s.\ \lambda (x, y).\ (x + 1, y)$ and $t_{y-1} = \lambda s.\ \lambda
  (x, y).\ (x, y - 1)$. Let $I = \setst{(x, y)}{xy \leq 0}$.
  %
  $\setst{(x, y)}{\sTIreachable{(x, y)}} = \nats \times -\nats \subseteq I$, so
  $O$ is \sTIconfluent. However, let $s_1 = (1, -1)$ and $s_2 = (0, 1)$.
  $I(s_1)$, $I(s_2)$, and $I(t_{x+1}(s_1)(s_1))$. However, $t_{x+1}(s_1)(s_2) =
  (1, 1)$, and $\lnot I((1, 1))$. Thus, $O$ is not \Iclosed{}.
\end{example}

\input{figs/not_necessary.fig}

Let's take a closer look at \exampleref{StateBasedNotNecessary} (illustrated in
\figref{StateBasedNotNecessary}) to understand why being \sTIconfluent{}
doesn't imply being \Iclosed{}. Our counterexample considers points $s_1 = (-1,
1)$ and $s_2 = (1, -1)$. $s_2$ is reachable, but $s_1$ is not. This is not a
coincidence! If we consider an \sTIconfluent{} object and set of transactions,
then any two reachable points satisfy the invariant, and the join of the two
points is again reachable, so it must satisfy the invariant as well. The only
way to join two invariant satisfying points and end up \emph{not} satisfying
the invariant is if one or both points are not reachable. Thus, if we know that
$I$ is a subset of reachable points, then being \sTIconfluent{} and being
\Iclosed{} are equivalent.

\begin{claim}\clmlabel{SufficientAndNecessary}
  If every invariant satisfying state is reachable, then being \sTIconfluent{}
  and being \Iclosed{} are equivalent.
  \[
    I \subseteq \setst{s}{\sTIreachable{s}} \implies
    (I(s_0) \land \text{$O$ \Iclosed{}} \iff \text{$O$ \sTIconfluent})
  \]
\end{claim}
\begin{elidableproof}
  Consider a state-based object $O = (S, \join)$, a start state $s_0$, a set of
  transactions $T$, and an invariant $I$ where $I \subseteq
  \setst{s}{\sTIreachable{s}}$.
  %
  If $I(s_0)$ and $O$ is \Iclosed{}, then $O$ is \sTIconfluent{} whether or not
  $I \subseteq \setst{s}{\sTIreachable{s}}$. Thus, we only have to prove that
  if $O$ is \sTIconfluent{}, then $I(s_0)$ and $O$ is \Iclosed.
  %
  If $O$ is \sTIconfluent{}, then $I(s_0)$. Next, let $s_1$ and $s_2$ be
  arbitrary states such that $I(s_1)$ and $I(s_2)$. Both states satisfy the
  invariant, so both are reachable. That is, there exist expressions $e_1$ and
  $e_2$ where $eval(e_1) = s_1$  and $eval(e_2) = s_2$. $e_1 \join e_2$ is
  again reachable, so since $T$ is \sTIconfluent{}, $I(s_1 \join s_2)$.  Thus,
  $O$ is \Iclosed.
\end{elidableproof}

Here are some more examples. \todo{Think of more compelling examples.}

\begin{example}
  Imagine that we have a set $P$ of students ($P$ is for pupils) and that we
  are trying to assign students into groups. We have a state based object $O =
  (S, \join)$ where each state $s \in S$ in a dictionary mapping a group id $1,
  \ldots, n$ to some subset of $P$. For example, if $P = \set{a, b, c}$, then
  $s$ might be $\set{1 \mapsto \set{a, b}, 2 \mapsto \set{c}}$. We join two
  dictionaries by taking a group-wise union. That is, $(s_1 \join s_2)(i) =
  s_1(i) \cup s_2(i)$.

  Furthermore assume each student $p$ has a year $year(p)$ (e.g. 5th grade, 4th
  grade). Our invariant $I$ is that every group of students is in the same
  year. That is, for every state $s$ and every $1 \leq i \leq n$, every student
  in $s(i)$ belongs to the same year.
  %
  Our set $T$ of transactions includes transactions which allow us to add any
  number of students to any number of groups.
  %
  Let $P = \set{a, b, \ldots, z}$ and $n = 3$. Our start state is $s_0 = \set{1
  \mapsto \set{a}, 2 \mapsto \set{b}, 3 \mapsto \set{c}}$.

  Some thought shows that O is \sTIconfluent{}. However, $O$ is not \Iclosed{}.
  Let
  \[
    s_1 = \set{1 \mapsto \set{x}, 2 \mapsto \emptyset, 3 \mapsto \emptyset}
    \qquad
    s_2 = \set{1 \mapsto \set{y}, 2 \mapsto \emptyset, 3 \mapsto \emptyset}
  \]
  where $year(x) \neq year(y)$. $I(s_1)$ and $I(s_2)$, but $\lnot I(s_1 \join
  s_2)$.
\end{example}

\begin{example}
  Consider the state-based object $O = (\powerset{\nats}, \cup)$ and start
  state $s_0 = \set{0}$. Let $T = \setst{\lambda X.\ X \cup Y}{Y \subseteq
  \nats}$ let $I = \setst{X}{\text{$X$ contains only odd or only even
  elements}}$. The $\sTIreachable{}$ states are $\powerset{2\nats}$, all of
  which satisfy $I$, so $O$ is \sTIconfluent{}.  However, let $A = \set{0}$ and
  $Y = \set{1}$. $I(A)$ and $I(B)$, but $A \join B = \set{0, 1}$ does not
  satisfy $I$.
\end{example}

\begin{example}
  Consider a state-based object consisting of two integer 2P-sets $(A_1, R_1)$
  and $(A_2, R_2)$ with initial state $(\set{1, 3}, \emptyset{})$ and $(\set{1,
  2, 3, 4}, \emptyset{})$. Let $T$ consist of all transactions that either
  remove an element from the first 2P-set (i.e.\ add something to $R_1$) or add
  something to the second 2P-set (i.e.\ add something to $A_2$). Let $I$ be
  that the first 2P-set is a subset of the second (i.e. $(A_1 - R_1) \subseteq
  (A_2 - R_2)$).
  %
  $O$ is \sTIconfluent{}. However, let $s_1 = (\set{1}, \set{2}, \set{1},
  \set{2})$ and $s_2 = (\set{2}, \set{1}, \set{2}, \set{1})$. $I(s_1)$ and
  $I(s_2)$, but $s_1 \cup s_2 = (\set{1, 2}, \set{1, 2}, \set{1, 2}, \set{1,
  2})$ which does not satisfy $I$.
\end{example}

\begin{example}
  Here are some of the example objects and invariants
  in~\cite{bailis2014coordination}. We begin with invariants over G-Sets $A$
  and $B$. Fortunately, a good number of invariants are \Iclosed{}.

  \begin{center}
    \begin{tabular}{|l|l|p{3in}|}
      \hline
      $I$ & \Iclosed{} & Description \\\hline\hline
      %
      $\forall x \in A.\ \phi(x)$ &
      yes &
        Invariants of this form include things like attribute equality and
        attribute inequality.
        \\\hline
      $\exists x \in A.\ \phi(x)$ &
      yes &
        Invariants of this form include things like non-emptiness constraints.
        \\\hline
      $\forall x \in A.\ \forall y \in A.\ \phi(x, y)$ &
      no &
        Invariants of this form include things like uniqueness constraints
        (e.g. $\forall x \in A.\ \forall y \in A.\ x \neq y \implies key(x)
        \neq key(y)$).
        \\\hline
      $\forall x \in A.\ \forall y \in B.\ \phi(x, y)$ &
      no &
        Invariants of this form include things like disjointness (e.g. $\forall
        x \in A. \forall y \in B.\ x \neq y$).
        \\\hline
      $\forall x \in A.\ \exists y \in A.\ \phi(x, y)$ &
      yes &
        Invariants of this form include things like foreign key constraints
        into a single relation. For example, if we have an \texttt{Employee(id,
        bossid)} relation, we might want the invariant $\forall (i_1, b_1) \in
        E.\ \exists (i_2, b_2) \in E.\ b_1 \neq \bot \implies b_1 = i_2$.
        \\\hline
      $\forall x \in A.\ \exists y \in B.\ \phi(x, y)$ &
      yes &
        Invariants of this form include things like foreign key constraints.
        \\\hline
      $\exists x \in A.\ \forall y \in A.\ \phi(x, y)$ &
      no &
        \todo{Figure out useful interpretation of these formula.}
        \\\hline
      $\exists x \in A.\ \forall y \in B.\ \phi(x, y)$ &
      no &
        \todo{Figure out useful interpretation of these formula.}
        \\\hline
      $\exists x \in A.\ \exists y \in A.\ \phi(x, y)$ &
      yes &
        \todo{Figure out useful interpretation of these formula.}
        \\\hline
      $\exists x \in A.\ \exists y \in B.\ \phi(x, y)$ &
      yes &
        \todo{Figure out useful interpretation of these formula.}
        \\\hline
    \end{tabular}
  \end{center}

  For 2P-Sets $A$ and $B$, we see that things get a bit more grim.
  \begin{center}
    \begin{tabular}{|l|l|}
      \hline
      $I$ & \Iclosed{} \\\hline\hline
      %
      $\forall x \in A.\ \phi(x)$ & yes \\\hline
      $\exists x \in A.\ \phi(x)$ & no \\\hline
      $\forall x \in A.\ \forall y \in A.\ \phi(x, y)$ & no \\\hline
      $\forall x \in A.\ \forall y \in B.\ \phi(x, y)$ & no \\\hline
      $\forall x \in A.\ \exists y \in A.\ \phi(x, y)$ & no \\\hline
      $\forall x \in A.\ \exists y \in B.\ \phi(x, y)$ & no \\\hline
      $\exists x \in A.\ \forall y \in A.\ \phi(x, y)$ & no \\\hline
      $\exists x \in A.\ \forall y \in B.\ \phi(x, y)$ & no \\\hline
      $\exists x \in A.\ \exists y \in A.\ \phi(x, y)$ & no \\\hline
      $\exists x \in A.\ \exists y \in B.\ \phi(x, y)$ & no \\\hline
    \end{tabular}
  \end{center}
  \note{%
    This would be a good table to demonstrate that \Iclosure{} is not only
    unnecessary, but also uncommon. In practice, not a lot of things are going
    to be \Iclosed{}. This would motivate our push beyond these simple checks.
    But then again, not a lot of things are going to be invariant-confluent
    either, so it's not a super strong argument.
  }
\end{example}
